---
title: "Bootstrapping"
author: "Math 271"
date: "3/10/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(magrittr)
library(moderndive)
library(purrr)
library(ggplot2)

```


```{r}
bind_group <- function(data, .id="run"){
  data %>% bind_rows(.id=.id) %>% group_by(.data[[.id]], .add=TRUE)
}
```

Working through chaper 8: https://moderndive.com/8-confidence-intervals.html

## Exploration

From `moderndive` package, `pennies_sample` is contents of a roll of pennies from a bank 
    + 50 pennies
    + an `ID` number and a `year` minted

```{r}
pennies_sample
pennies_sample %>% summarize(year=mean(year))
ggplot(pennies_sample) + aes(x=year) + geom_histogram(center=0, binwidth=5)
```

Simulate getting new samples? (Without getting more pennies...)

Make many "copies" of your sample. Sample from copies.

```{r}
pseudo_population <- rerun(100, pennies_sample) %>% bind_rows
pseudo_population %>% bind_rows %>% slice_sample(n=50) %>% summarize(mean(year))
```

#Sampling With Replacement Method 

Or: sample __with replacement__ from your original sample.

Same as creating "infinitely" many copies in previous idea.

```{r}
pennies_sample %>% slice_sample(prop=1, replace=TRUE) %>% summarize(mean(year))
```

See effect of sampling on `mean(year)`, get many resamples.

This example reruns 1000 bootstrap samples:

```{r}
penny_means <- rerun(1000, pennies_sample %>% slice_sample(prop=1, replace=TRUE)) %>% 
  bind_group(.id="sample") %>% 
  summarize(mean_year= mean(year))

ggplot(penny_means) + aes(x=mean_year) + geom_histogram(center=0, binwidth=1) + 
  xlim(1985, 2005) + geom_vline(xintercept=1995, color="red") + 
  geom_vline(xintercept=c(1992, 1999), col="blue")

```
Numerical summaries of estimated sampling distribution for `mean(year)`.

Also, quantiles from the sampling distribution

Quantiles from matching normal distribution

```{r}
(penny_ci <- penny_means %>% summarize(min=min(mean_year), max=max(mean_year), 
                                       mean=mean(mean_year), sd=sd(mean_year), 
                                       level=0.95, 
                                       p = c( (1-level)/2, (1+level)/2),
                                       perc = quantile(mean_year, p), 
                                       norm = qnorm(p, mean, sd)))
```
Use as __confidence intervals__ 
    + "confidence" is `level` that the mean of population is inside
    + "confidence" = This method will produce a "correct" interval  with at least `level` chance
    + "correct" = the population mean is inside

```{r}
ggplot(penny_means) + aes(x=mean_year) + geom_histogram(aes(y=..density..), center=0, binwidth=1) + 
  xlim(1985, 2005) + geom_vline(xintercept=1995, color="red") + 
  geom_vline(xintercept=c(1992, 1999), col="blue") + geom_density() + 
  #geom_function(fun=function(x) dnorm(x, mean=1995.49, sd=2.22), col="purple")
  geom_function(fun=dnorm, args=list(mean=1995.49, sd=2.22), col="orange")

ggplot(penny_means) + aes(x=mean_year) + geom_histogram(aes(y=..density..), center=0, binwidth=1) + 
  geom_function(fun=dnorm, args=list(mean=1995.49, sd=2.22), col="orange") + 
  geom_vline(aes(xintercept=value, color=type), data=penny_ci %>% pivot_longer(perc|norm, names_to="type"))

```

